---
title: "EC3 Elements Report Analysis"
#author: "Eden Axelrad"
output: 
  html_document:
    theme: paper
---

<style>

.nav-tabs > li.active > a,
.nav-tabs > li.active > a:hover,
.nav-tabs > li.active > a:focus,
.nav-tabs > li.active > a:focus:hover {
    webkit-box-shadow: inset 0 -2px 0 #398D93;
    box-shadow: inset 0 -2px 0 #398D93;
    color: #398D93;
}

.nav-tabs > li > a:hover,
.nav-tabs > li > a:focus,
.nav-tabs > li > a:focus:hover {
    box-shadow: inset 0 -2px 0 #87C1DE;
    color: #87C1DE;
}

</style>

```{r setup, include=FALSE}

# Load required libraries
library(tidyverse)
library(readxl)
library(highcharter)
library(plotly)
library(DT)
library(knitr)

```

```{r get data, echo=FALSE, warning=FALSE, include=FALSE}

# Read in the Master format table
Masterformat_Codes <- read.csv("Data/Masterformat 2 Codes.csv")

# Cleanup function that takes a single report and provides an organized data frame
File_Path <- paste0("Element Reports/", "00 CLF LA Example-elements-report.xlsx")

# Retrieve project header info (name, level of dev, area, date of report, and data source)
Project_Info <- c(
  read_excel(File_Path, sheet = "Sheet2", col_names = FALSE)[1:3, 1:2]  %>% deframe(),
  read_excel(File_Path, sheet = "Sheet2", col_names = FALSE)[1:2, 4:5]  %>% deframe()) 

# Create parameters from the area and project name 
Area <- as.numeric(Project_Info["Area"])
Project <- Project_Info["EC3 Building Project"]

# Read in and clean the report
Report <- read_excel(File_Path, sheet = "Sheet2", skip = 4) %>%
  # Remove any rows that have an A5 sub-assembly (construction)
  filter(Subassembly != "A5 Construction") %>%
  rename("Baseline (kgCO2e)" = 7,
         "Conservative (kgCO2e)" = 8,
         "Realized (kgCO2e)" = 9,
         "Achievable (kgCO2e)" = 11,
         "Selected Material Name" = 15,
         "Selected Material Manufacturer" = 13,
         "Selected Material Plant" = 14) %>%
  # Add project name, area, reduction from baseline and master format codes
  mutate(Project = Project, 
         Area = Area,
         Unit = gsub("1 ", "", unit),
         `Reduction from Baseline` = 1-(`Realized (kgCO2e)`/`Baseline (kgCO2e)`),
         `Masterformat Code` = substr(Masterformat, 1, 8),
         `M1 Code` = substr(`Masterformat Code`, 1, 2),
         `M2 Code` = str_squish(substr(`Masterformat Code`, 1, 5)),
         `Masterformat 3` = ifelse(substr(`Masterformat Code`, 7, 8) == "00", 
                                   "", str_squish(gsub('[0-9]+', '', Masterformat)))) %>%
  mutate_if(is.numeric, round, digits = 4) %>%
  # Merge with named lists to replace codes with actual names
  left_join(
    Masterformat_Codes %>%
      filter(Level == "Level 1") %>%
      mutate(`M1 Code` = substr(L2.Code, 1, 2)) %>%
      select(`M1 Code`, "Masterformat 1" = Name),
    by = "M1 Code") %>%
  left_join(
    Masterformat_Codes %>%
      mutate(`M2 Code` = str_squish(substr(L2.Code, 1, 6))) %>%
      select(`M2 Code`, "Masterformat 2" = Name, Assembly),
    by = "M2 Code") %>%
  # Reorder columns and drop non-relevant columns
  select(Project, Area, `Masterformat Code`, `Masterformat 1`, `Masterformat 2`, `Masterformat 3`, 
         Assembly, Element, Material, Quantity, Unit, `Baseline (kgCO2e)`, `Conservative (kgCO2e)`, 
         `Realized (kgCO2e)`, `Achievable (kgCO2e)`, `Reduction from Baseline`,
         `Selected Material Name`, `Selected Material Manufacturer`, `Selected Material Plant`) %>%
  drop_na(Quantity)

```

<br>

## Results for `r return(Project)`{.tabset}

<br>

### Project Info {.tabset}

<br>

```{r overall, echo=FALSE, fig.align = 'center'}

Project_Info <- Project_Info %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  rename("Parameter" = 1, "Value" = 2) %>%
  rbind.data.frame(
    c("Total Embodied Carbon (kgCO2e)", 
      scales::comma(sum(Report$`Realized (kgCO2e)`))),
    c("EC Intensity (kgCO2e/sqft)", 
      round(sum(Report$`Realized (kgCO2e)`)/Area, digits = 2)),
    c("Reduction from Baselline (kgCO2e)", 
      scales::comma(sum(Report$`Baseline (kgCO2e)`)-sum(Report$`Realized (kgCO2e)`))),
    c("Reduction from Baseline (%)", 
      scales::percent((sum(Report$`Baseline (kgCO2e)`)-sum(Report$`Realized (kgCO2e)`))/sum(Report$`Baseline (kgCO2e)`), 
                      accuracy = 0.01))) 

Project_Info[3, 1:2] <- c("Area (square feet)", scales::comma(Area))

Project_Info %>%
  kable() 

```

<br>

```{r scenario, echo=FALSE, warning=FALSE, fig.align = 'center'}

# Create a table that has total EC by scenario
By_Scenario <- Report %>%
  select(`Masterformat 1`, `Baseline (kgCO2e)`, `Conservative (kgCO2e)`,  `Achievable (kgCO2e)`, `Realized (kgCO2e)`) %>%
  rename("CSI Division" = 1, "Baseline" = 2, "Conservative" = 3, "Achievable" = 4, "Realized" = 5) %>%
  pivot_longer(-`CSI Division`,
               names_to = "Scenario",
               values_to = "Embodied Carbon (kgCO2e)") %>%
  group_by(Scenario) %>%
  summarize("Embodied Carbon (mtCO2e)" = sum(`Embodied Carbon (kgCO2e)`)/1000) %>%
  mutate(Scenario = factor(Scenario, levels = c("Baseline", "Conservative", 
                                                "Achievable", "Realized"))) %>%
  arrange(Scenario)

# Generate a plot showing the total EC by scenario
By_Scenario %>%
  ggplot(aes(x = Scenario, y = `Embodied Carbon (mtCO2e)`, fill = Scenario)) +
  geom_col(width = .8) +
  scale_fill_manual(values = c("#424242", "#9A9795", "#C8C8C8", "#398D93")) +
  scale_y_continuous(labels = scales::comma) +
  scale_x_discrete(labels = c("Realized" = expression(bold(Realized)), parse = TRUE)) +
  geom_segment(aes(x = 0.6, 
                   xend = 4.4, 
                   y = By_Scenario$`Embodied Carbon (mtCO2e)`[4], 
                   yend = By_Scenario$`Embodied Carbon (mtCO2e)`[4]),
               linetype = "dashed", 
               color = "#D73833") +
  geom_text(aes(x = 1, y = `Embodied Carbon (mtCO2e)`[1], 
                label = scales::comma(`Embodied Carbon (mtCO2e)`[1])), 
            vjust = -.5, size = 3, fontface = "plain") +
  geom_text(aes(x = 2, y = `Embodied Carbon (mtCO2e)`[2], 
                label = scales::comma(`Embodied Carbon (mtCO2e)`[2])), 
            vjust = -.5, size = 3, fontface = "plain") +
  geom_text(aes(x = 3, y = `Embodied Carbon (mtCO2e)`[3], 
                label = scales::comma(`Embodied Carbon (mtCO2e)`[3])), 
            vjust = -.5, size = 3, fontface = "plain") +
  geom_text(aes(x = 4, y = `Embodied Carbon (mtCO2e)`[4], 
                label = scales::comma(`Embodied Carbon (mtCO2e)`[4])), 
            vjust = -.5, size = 3.5, fontface = "bold") +
  labs(title = "Embodied Carbon by Scenario",
       x = "") +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        legend.position = "none",
        text = element_text(family = "Arial"))

```


<br>

### By Division {.tabset}

```{r plot function, warning=FALSE, include=FALSE}

# summary table needed for the plot function
Summary_Table <- Report %>%
  group_by(`Masterformat 1`) %>%
  summarize(Baseline = sum(`Baseline (kgCO2e)`), 
            Realized = sum(`Realized (kgCO2e)`)) %>%
  pivot_longer(!`Masterformat 1`, names_to = "Type", values_to = "kgCO2e") %>%
  left_join(
    Report %>%
      group_by(`Masterformat 1`) %>%
      summarize(Baseline = sum(`Baseline (kgCO2e)`), 
                Realized = sum(`Realized (kgCO2e)`)) %>%
      mutate(Change = Realized-Baseline,
             `Percent Change` = Change/Baseline),
    by = "Masterformat 1") %>%
  mutate(Change = ifelse(Type == "Realized", Change, NA),
         `Percent Change` = ifelse(Type == "Realized", `Percent Change`, NA)) 

# Plot for individual categories
Plot_Mstfmt1_Comp <- function(Mstfmt) {
  
  baseline <- pluck(Summary_Table %>% 
                      filter(`Masterformat 1` == Mstfmt) %>% 
                      .$Baseline, 2)
  
  realized <- pluck(Summary_Table %>% 
                      filter(`Masterformat 1` == Mstfmt) %>% 
                      .$Realized, 2)
  
  sub_txt <- paste0(
    pluck(Summary_Table %>% 
            filter(`Masterformat 1` == Mstfmt) %>% 
            .$`Percent Change`, 2) %>%
      abs() %>%
      scales::percent(accuracy = 0.1), 
    ifelse(realized > baseline, " Increase", " Reduction"), 
    " (", scales::comma(realized-baseline), " kgCO2e)")
  
  Plot <- Summary_Table %>%
    filter(`Masterformat 1` == Mstfmt) %>%
    ggplot(aes(x = Type, y = kgCO2e, fill = Type)) +
    geom_col(width = .7) +
    scale_fill_manual(values = c("#424242", "#398D93")) +
    geom_segment(aes(x = 0.65, xend = 2.35, y = baseline, yend = baseline),
                 linetype = "dashed", linewidth = 0.75, color = "black") +
    {if(realized != baseline)
      geom_segment(aes(x = 2, 
                       y = baseline, 
                       xend = 2, 
                       yend = realized),
                   lineend = "round",
                   linejoin = "bevel",
                   arrow = arrow(length = unit(0.2, "cm")),
                   linewidth = 0.5,
                   color = "black")}+
    scale_y_continuous(labels = scales::comma) +
    labs(title = Mstfmt,
         subtitle = sub_txt,
         x = "") +
    theme_minimal() +
    theme(panel.grid.major.x = element_blank(),
          legend.position = "none")
  
  Plot
  
}

```

```{r summary table, echo=FALSE, fig.align = 'center'}

Report %>%
  mutate(`Masterformat Division` = paste(substr(`Masterformat Code`, 1, 2), `Masterformat 1`)) %>%
  group_by(`Masterformat Division`) %>%
  summarize(Baseline = sum(`Baseline (kgCO2e)`), 
            Realized = sum(`Realized (kgCO2e)`)) %>%
  mutate(Change = Realized-Baseline,
         `Percent Change` = scales::percent(Change/Baseline)) %>%
  mutate_if(is.numeric, scales::comma) %>%
  kable()

```

<br>

```{r dynamic tab, echo=FALSE, results='asis', fig.align = 'center'}

tabs <- unique(Report$`Masterformat 1`)

for(tab in tabs) {

  cat('\n')
  cat('#### ', tab, '   \n')
  cat('\n')

  plot_values <- Report %>% filter(`Masterformat 1` == tab)

  print(Plot_Mstfmt1_Comp(tab))

  cat('\n')

}

```

### By Assembly

```{r assembly table, echo=FALSE, fig.align = 'center'}

Report %>%
  group_by(Assembly) %>%
  summarize(Baseline = sum(`Baseline (kgCO2e)`), 
            Realized = sum(`Realized (kgCO2e)`)) %>%
  mutate(Change = Baseline-Realized,
         `Percent Change` = scales::percent(Change/Baseline)) %>%
  mutate_if(is.numeric, scales::comma) %>%
  kable()

```

<br>

```{r assembly chart, echo=FALSE, fig.align = 'center'}

Report %>%
  group_by(Assembly) %>%
  summarize(Baseline = sum(`Baseline (kgCO2e)`), 
            Realized = sum(`Realized (kgCO2e)`),
            Achievable = sum(`Achievable (kgCO2e)`)) %>%
  pivot_longer(cols = c(Baseline, Realized, Achievable),
               names_to = "Calculation",
               values_to = "kgCO2e") %>%
  mutate(Calculation = factor(Calculation, levels = c("Baseline", "Realized", "Achievable"))) %>%
  ggplot(aes(x = Calculation, y = kgCO2e, fill = Calculation)) +
  geom_col() +
  scale_fill_manual(values = c("#424242", "#398D93", "#C8C8C8")) +
  scale_y_continuous(labels = scales::label_comma()) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  facet_wrap(~Assembly, scales = "free") +
  labs(title = "Realized vs Baseline and Achievable",
       subtitle = "By Assembly",
       x = "") +
    theme_minimal() +
    theme(panel.grid.major.x = element_blank(),
          legend.position = "none")

```

### Bar Chart

<br>

```{r stacked bar plotly, include=FALSE, echo=FALSE, warning=FALSE, out.width="100%"}

# Color palette
colors <- c("#9A9795", "#EFC143", "#6AB1D5", "#2A378E", "#F04A52", "#F5C4D3", "#4D9A44", "#276168")

# Create a stacked bar chart at 3 levels of detail - total, master 1, and master 2
stacked_plot <- rbind.data.frame(
  Report %>%
    mutate(Group = paste0(substr(`Masterformat Code`, 1, 2), " ", `Masterformat 1`, ": ", `Masterformat 2`)) %>%
    group_by(Group) %>%
    summarize(`Embodied Carbon (kgCO2e)` = sum(`Realized (kgCO2e)`)) %>%
    mutate(Level = "Masterformat 2"),
  Report %>%
    mutate(Group = paste(substr(`Masterformat Code`, 1, 2), `Masterformat 1`)) %>%
    group_by(Group) %>%
    summarize(`Embodied Carbon (kgCO2e)` = sum(`Realized (kgCO2e)`)) %>%
    mutate(Level = "Masterformat 1"),
  data.frame(
    Group = "Total",
    Value = sum(Report$`Realized (kgCO2e)`),
    Level = "Total") %>%
    rename("Group" = Group, "Embodied Carbon (kgCO2e)" = Value, "Level" = Level)) %>%
  mutate(Level = factor(Level, levels = c("Total", "Masterformat 1", "Masterformat 2")),
         fill_grp = case_when(
           Level == "Total" ~ "Total",
           Level == "Masterformat 1" ~ Group,
           Level == "Masterformat 2" ~ sub(":.*", "", Group))) %>%
  ggplot(aes(x = Level, y = `Embodied Carbon (kgCO2e)`, fill = fill_grp,
             text = paste0("<b>", Group, "</b>", 
                           "<br>Emissions (kgCO2e): ", 
                           scales::comma(`Embodied Carbon (kgCO2e)`),
                           "<br>Share of Total: ", 
                           scales::percent(`Embodied Carbon (kgCO2e)`/sum(Report$`Realized (kgCO2e)`), 
                                           accuracy = 0.01)))) +
  geom_bar(position = "stack", stat = "identity", color = "white", linewidth = .3) +
  scale_fill_manual(values = colors) +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(x = "",
       fill = "Division",
       title = "Total Embodied Carbon by Level of Detail") +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank())

# Make it interactive with ggplotly 
ggplotly(stacked_plot, tooltip = "text") %>%
  config(displayModeBar = FALSE)

```

```{r stacked bar highcharter, echo=FALSE, warning=FALSE, out.width="100%"}

# Load the hicharter library
library(highcharter)

# Create a color palette for the bar chart
colors <- c("#9A9795", "#EFC143", "#6AB1D5", "#2A378E", "#F04A52", "#F5C4D3", "#4D9A44", "#276168")

# create the same df as above but add 2 labels columns for emissions and share of total
stacked_df <- rbind.data.frame(
  Report %>%
    mutate(Group = paste0(substr(`Masterformat Code`, 1, 2), " ", `Masterformat 1`, ": ", `Masterformat 2`)) %>%
    group_by(Group) %>%
    summarize(`Embodied Carbon (kgCO2e)` = sum(`Realized (kgCO2e)`)) %>%
    mutate(Level = "Masterformat 2"),
  Report %>%
    mutate(Group = paste(substr(`Masterformat Code`, 1, 2), `Masterformat 1`)) %>%
    group_by(Group) %>%
    summarize(`Embodied Carbon (kgCO2e)` = sum(`Realized (kgCO2e)`)) %>%
    mutate(Level = "Masterformat 1"),
  data.frame(
    Group = "Total",
    Value = sum(Report$`Realized (kgCO2e)`),
    Level = "Total") %>%
    rename("Group" = Group, "Embodied Carbon (kgCO2e)" = Value, "Level" = Level)) %>%
  mutate(Level = factor(Level, levels = c("Total", "Masterformat 1", "Masterformat 2")),
         fill_grp = case_when(
           Level == "Total" ~ "Total",
           Level == "Masterformat 1" ~ Group,
           Level == "Masterformat 2" ~ sub(":.*", "", Group)),
         EC_label = scales::comma(`Embodied Carbon (kgCO2e)`),
         pct_label = scales::percent(`Embodied Carbon (kgCO2e)`/sum(Report$`Realized (kgCO2e)`), 
                                           accuracy = 0.01))

# Create the chart, identical to the previous plotly one
stacked_df %>%
  hchart(type = 'column', 
         stacking = 'normal',
         hcaes(x = Level, y = `Embodied Carbon (kgCO2e)`, group = fill_grp)) %>%
  hc_legend(align = "right", verticalAlign = "middle", layout = "vertical") %>%
  hc_colors(colors = colors) %>%
  hc_title(text = "Embodied Carbon by Level",
           style = list(fontWeight = "bold")) %>% 
  hc_subtitle(text = "A breakdown of embodied carbon by masterformat level 1 and 2") %>%
  hc_xAxis(title = "", categories = levels(stacked_df$Level)) %>%
  hc_tooltip(valueDecimals = 0,
             pointFormat = "<b>{point.Group}</b> <br> 
             Emissions (kgCO2e): {point.EC_label} <br> 
             Share of Total: {point.pct_label}") %>%
  hc_add_theme(hc_theme_smpl())

```



### Waterfall Chart

<br>

```{r waterfall, echo=FALSE, warning=FALSE, out.width="80%"}

# Create a data frame for the waterfall chart 
wtrfl_df <- rbind.data.frame(
  data.frame(
    Label = "Baseline (mtCO2e)",
    Change = round(sum(Report$`Baseline (kgCO2e)`)/1000, digits = 2)), 
  Report %>%
    mutate(Label = paste0(substr(`Masterformat Code`, 1, 2), " ", `Masterformat 1`, ": ", `Masterformat 2`)) %>%
    group_by(Label) %>%
    summarize(Baseline = sum(`Baseline (kgCO2e)`), 
              Realized = sum(`Realized (kgCO2e)`)) %>%
    mutate(Change = round((Realized-Baseline)/1000, digits = 2)) %>%
    filter(Change != 0) %>%
    select(Label, Change)) %>%
  mutate(Cumulative = cumsum(Change)) %>%
  rbind.data.frame(
    data.frame(
      Label = "Realized (mtCO2e)",
      Change = round(sum(Report$`Realized (kgCO2e)`)/1000, digits = 2),
      Cumulative = round(sum(Report$`Realized (kgCO2e)`)/1000, digits = 2))) %>%
  mutate(Order = rev(seq(1:nrow(.))),
         Label = fct_reorder(Label, Order),
         Start = ifelse(Label %in% c("Baseline (mtCO2e)", "Realized (mtCO2e)"), 0, Cumulative-Change),
         Color = case_when(
           Label == "Baseline (mtCO2e)" ~ "#424242",
           Label == "Realized (mtCO2e)" ~ "#398D93",
           Change < 0 ~ "#33b86f",
           Change > 0 ~ "#c94426"))


# Use the above data frame to generate a ggplot waterfall chart
wtrfl_df %>%
  ggplot(aes(fill = Color)) +
  geom_rect(aes(x = Label, 
                xmin = Order - 0.45, xmax = Order + 0.45, 
                ymin = Cumulative, ymax = Start),
            color = "black") +
  scale_fill_identity() +
  scale_y_continuous(label = scales::comma) +
  geom_hline(yintercept = 0) +
  labs(x = "", y = "mtCO2e",
       title = "Waterfall Chart: Realized vs Baseline",
       subtitle = "(only shows categories with a change from baseline)") +
  geom_text(aes(x = Label, 
                y = wtrfl_df$Change[nrow(wtrfl_df)]*.9, 
                label = scales::comma(Change, digits = 1)), 
            hjust = 1, 
            fontface = ifelse(wtrfl_df$Label %in% c("Baseline (mtCO2e)", "Realized (mtCO2e)"), "bold", "plain"), 
            color = ifelse(wtrfl_df$Label %in% c("Baseline (mtCO2e)", "Realized (mtCO2e)"), "white", 
                           ifelse(wtrfl_df$Change < 0, "#33b86f", "#c94426")),) +
  coord_flip() +
  theme(legend.position = "none",
        axis.ticks = element_blank(),
        plot.subtitle = element_text(size = 9, color = "#424242"))

```

### EC3 Table

```{r table, echo=FALSE, warning=FALSE}

datatable(Report[, c(4:8, 10:12, 14, 16, 17)],
              caption = paste0("A full table of elements for ", Project),
              class = 'cell-border stripe',
              rownames = FALSE,
              filter = 'top',
              options = list(dom = 'tip',
                             columnDefs = list(list(className = 'dt-left', targets = 0:10)))) %>%
  formatStyle(columns = c("Element", "Selected Material Name"), fontSize = '75%') %>%
  formatPercentage(columns = c("Reduction from Baseline"), digits = 1) %>%
  formatRound(columns = c("Quantity", "Baseline (kgCO2e)", "Realized (kgCO2e)"), digits = 0)

```
